---
title: "Frog Occurences Australia"
---


## TidyTuesday Section

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

## Research Question: How are frog species occurrences distributed across Australian states?”

```{r}
# Using R
# Option 1: tidytuesdayR R package 
# install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2025-09-02')
## OR
#tuesdata <- tidytuesdayR::tt_load(2025, week = 35)

frogID_data <- tuesdata$frogID_data
frog_names <- tuesdata$frog_names

# Option 2: Read directly from GitHub

frogID_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-02/frogID_data.csv')
frog_names <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-02/frog_names.csv')
```
```{r}
#View(frogID_data)
```


```{r}
library(ggplot2)

ggplot(frogID_data, aes(x = stateProvince, fill = stateProvince)) +
  geom_bar() +
  labs(title = "Number of Frog Occurrences by State",
       x = "State/Province", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))


```
> Looking at the bar chart of frog occurrences by state, it is clear that frog records are unevenly distributed across Australia. New South Wales has by far the highest number of occurrences, with close to 60,000 records, followed by Victoria and Queensland (32000 and 21000 respectively) , which also show substantial numbers of observations. Other states, such as South Australia, Tasmania, and the Australian Capital Territory, contribute far fewer records, while the Northern Territory and Other Territories have very limited observations. 
