---
title: "Professional Viz"
---


```{r}
#packages
library(ggplot2)
library(dplyr)
library(readr)
library(plotly)
library(sf)
library(viridis)
```

### How did Crime fare by Province in South Africa from 2015 to 2016?

South Africa, with its diverse landscapes and complex social history, has long grappled with issues of crime that impact public safety, economic development, and community well-being.
This solo project aims to visualize the distribution of crime across South Africa's provinces during the 2015–2016 fiscal year using spatial data analysis and data visualization.\

\
**Objective:**

The primary objective of this research is to provide a clear, intuitive map-based representation of total reported crimes by province.

```{r}
# data
crime <- read_csv("../../data/raw/SouthAfricaCrimeStats_v2.csv")


# shape file
map_data <- st_read("../../data/raw/south_africa_Province_level_1_MAPOG/south_africa_Province_level_1.shp")


```

```{r}
# increase spread by log transformation
crime_filtered <- crime %>%
  select(Province, `2015-2016`) %>%
  mutate(`2015-2016` = as.numeric(`2015-2016`))
```

```{r}
# Inspect data
head(map_data)
head(crime)


```

```{r}
# group by province and create a sum for the entries
crime_filtered <- crime %>%
  group_by(Province) %>%
  summarise(`2015-2016` = sum(as.numeric(`2015-2016`))) 

# Standardize province names and lowercase & remove extra spaces
crime_filtered <- crime_filtered %>%
  mutate(Province = tolower(trimws(Province))) 

# correct KwaZulu-Natal Kwazulu/Natal mismatch
crime_filtered <- crime_filtered %>%
  mutate(Province = ifelse(Province == "kwazulu-natal", "kwazulu/natal", Province))

# rename shape1 to province
map_data <- map_data %>%
  rename(Province = shape1) %>%
  mutate(Province = tolower(trimws(Province)))
```

```{r}
# join the csv and the shape file
map_data <- map_data %>%
  left_join(crime_filtered, by = "Province")
```



```{r}
# plot
library(ggplot2)

ggplot() +
  geom_sf(data = map_data, aes(fill = `2015-2016`), color = "white", size = 0.15) +
  scale_fill_viridis_c(option = "plasma", na.value = "red") +
  labs(
    title = "Total Crime Cases in South Africa (2015-2016)",
    fill = "Cases",
    caption = "Source: South African Police Service (SAPS)"
  ) +
  theme_void()


```

#### Results

The resulting map visually emphasizes disparities in crime totals across provinces:

-   **Gauteng**, **Western Cape**, **Northen** and **KwaZulu-Nata**l (most of which are densely populated provinces) appear to have higher crime totals, possibly reflecting both population density and urban challenges.

-   On the other side of the spectrum we have **Free State**, **North West Province**, **Limpopo** and **Mpumalanga**; most of which are sparsely populated areas with relatively lower crime totals, though this does not necessarily imply safety, as underreporting and other factors might influence statistics
